---
title: "R Notebook"
output: html_notebook
---
First time running this code? Run "write_movebank_credentials.R" first.

```{r}
library(tidyverse)
library(sf)
library(move)
library(momentuHMM)
library(furrr)
```

```{r}
amwo_data <- getMovebankData(study = "American Woodcock Migration Ecology in Eastern North America", login = readRDS("movebank_credentials.rds"), removeDuplicatedTimestamps=TRUE, timestamp_end = "20220731000000000") %>% as.data.frame()

amwo_data <- amwo_data %>%
  transmute(ID = local_identifier,
            time = timestamp,
            date = strftime(timestamp, format="%Y-%m-%d"),
            x = location_long,
            y = location_lat,
            tagtype = comments,
            sex = sex,
            age = taxon_detail,
            altitude = height_above_msl,
            event_id = event_id)
```

Remove individuals with fewer than 3 locations, recording which ones were removed
```{r}
individuals_removed_for_too_few_locations <- amwo_data %>%
  distinct(ID, date, .keep_all = TRUE) %>% 
  group_by(ID) %>%
  tally() %>%
  filter(n < 3) %>%
  pull(ID)

birds_to_keep <- amwo_data %>%
  distinct(ID, date, .keep_all = TRUE) %>% 
  group_by(ID) %>%
  tally() %>%
  filter(n > 2) %>%
  pull(ID)

amwo_data <- amwo_data %>%
  filter(ID %in% birds_to_keep)
```

Trial code: reduce to a few individuals with known deaths to see if this works
Now trying with full subset
```{r}
amwo_data_subsampled <- amwo_data %>% 
  # filter(ID %in% c("VA-2019-24", "NJ-2019-26", "NJ-2019-25", "LA-2022-14", "NJ-2018-14", "SC-2019-09", "SC-2019-04")) %>% # Colby's identified deaths
  dplyr::select(ID, time, x, y, event_id) %>% 
  mutate(test = FALSE) %>% 
  group_by(ID) %>%
  mutate(ranks = row_number()) %>%
  mutate(known = NA) %>% 
  #mutate(known = ifelse(ranks == 1, 1, NA)) %>%
  ungroup() %>% 
  dplyr::select(-ranks)
```

Upload data from the ground test, which serves as a training dataset, and joining to the main dataset
```{r}
ground_test <- read_csv("ground_test_2_rformat.csv", 
                        col_types = cols(`GMT Time` = col_datetime(format = "%m/%d/%Y %H:%M"))) %>% 
  transmute(ID = as.character(id),
         time = `GMT Time`,
         x = Longitude,
         y = Latitude,
         event_id = 0,
         test = TRUE) %>% 
  group_by(ID) %>%
  mutate(ranks = row_number()) %>%
  mutate(known = 3) %>% 
  #mutate(known = ifelse(ranks == max(ranks), 3, NA)) %>% 
  ungroup() %>% 
  dplyr::select(-ranks)

amwo_data_subsampled <- bind_rows(ground_test, amwo_data_subsampled)
```

Determine the mean distance to the nearest 5 points
```{r}
# plan(sequential)
# 
# amwo_data_subsampled$mean_dist_5 <- amwo_data_subsampled %>% 
#   dplyr::select(ID, x, y) %>% 
#   future_pmap(function(ID_iter, x_iter, y_iter){
#     
#     iter_sf <- data.frame(x = x_iter, y = y_iter) %>% 
#       st_as_sf(coords = c("x", "y"), crs = 4326, remove = FALSE)
#     
#     other_locs <- amwo_data_subsampled %>%
#       st_as_sf(coords = c("x", "y"), crs = 4326, remove = FALSE) %>%
#       filter(ID == ID_iter)
#     
#     st_distance(x = iter_sf, y = other_locs) %>% 
#       as.numeric() %>%
#       sort() %>%
#       .[2:6] %>% 
#       mean(na.rm = TRUE) %>% 
#       return()
#   }) %>% 
#   unlist()
# 
# plan(sequential)
```

Determine the mean distance to the nearest 10 points
```{r}
# plan(sequential)
# 
# amwo_data_subsampled$mean_dist_10 <- amwo_data_subsampled %>% 
#   dplyr::select(ID, x, y) %>% 
#   future_pmap(function(ID_iter, x_iter, y_iter){
#     
#     iter_sf <- data.frame(x = x_iter, y = y_iter) %>% 
#       st_as_sf(coords = c("x", "y"), crs = 4326, remove = FALSE)
#     
#     other_locs <- amwo_data_subsampled %>%
#       st_as_sf(coords = c("x", "y"), crs = 4326, remove = FALSE) %>%
#       filter(ID == ID_iter)
#     
#     st_distance(x = iter_sf, y = other_locs) %>% 
#       as.numeric() %>%
#       sort() %>%
#       .[2:11] %>% 
#       mean(na.rm = TRUE) %>% 
#       return()
#   }) %>% 
#   unlist()
# 
# plan(sequential)
```

Determine the mean distance to the nearest 15 points
```{r}
plan(sequential)

amwo_data_subsampled$mean_dist_15 <- amwo_data_subsampled %>% 
  dplyr::select(ID, x, y) %>% 
  future_pmap(function(ID_iter, x_iter, y_iter){
    
    iter_sf <- data.frame(x = x_iter, y = y_iter) %>% 
      st_as_sf(coords = c("x", "y"), crs = 4326, remove = FALSE)
    
    other_locs <- amwo_data_subsampled %>%
      st_as_sf(coords = c("x", "y"), crs = 4326, remove = FALSE) %>%
      filter(ID == ID_iter)
    
    st_distance(x = iter_sf, y = other_locs) %>% 
      as.numeric() %>%
      sort() %>%
      .[2:16] %>% 
      mean(na.rm = TRUE) %>% 
      return()
  }) %>% 
  unlist()

plan(sequential)
```

```{r}
amwo_hmm <- amwo_data_subsampled %>% 
  as.data.frame() %>% 
  prepData(type = "LL")
```

```{r}
stateNames <- c("live mig", "live stop", "dead") #"live stop"
dist <- list(step="gamma", angle="wrpcauchy", mean_dist_15 = "norm") #mean_dist_5 = "norm", mean_dist_10 = "norm",

fixPar <- list(beta=matrix(c(NA, NA, NA, NA, -1000, -1000), nrow = 1, byrow = TRUE))
# 
Par0_m1 <- list(step=c(27.65267, 0.143942491, 0.01349499,
                       63.35321, 0.212567737, 0.01235371,
                       0.001, 0.002383503, 0.02842554),
                angle = c(3.1051583, 3.1136461, -3.0911715,
                          0.5042311, 0.2632081, 0.4294835),
                # mean_dist_5 = c(17216.08, 15.04497, 4.857182,
                #                 47633.62, 14.31838, 12.161531),
                # mean_dist_10 = c(17216.08, 15.04497, 4.857182,
                #                 47633.62, 14.31838, 12.161531),
                mean_dist_15 = c(33528.00, 21.98520, 6.911486,
                                86789.51, 19.83828, 13.351950))
ks <- amwo_hmm$known
```


```{r}
mod <- fitHMM(data=amwo_hmm, nbStates=3, dist=dist, Par0 = Par0_m1,
               estAngleMean = list(angle=TRUE),
               fixPar=fixPar,
               #userBounds = userBounds,
               #workBounds = workBounds,
               #DM = DM,
               knownStates = ks,
               #formula = ~step_cov, #state 2 to 3. page 47 of vignette shows how to create the beta matrix
               stateNames = stateNames)
  
mod_vit <- amwo_hmm %>% 
  mutate(point_state = viterbi(mod))

mod
```

Save results for supervised classification in the app 
```{r}
saveRDS(mod_vit, file = "predicted_mortalities.rds")
saveRDS(individuals_removed_for_too_few_locations, file = "birds_removed.rds")
```

